\documentclass[10pt]{reportMaster}

\title{Automated Prediction of Experimental Metadata from Scientific Publications}
\author{Stuti\ Nayak}
\id{DKE-19-06}
 \nameMastersProgramme{Data Science for Decision Making}  
%\nameMastersProgramme{Artificial Intelligence}
\committee{Dr. Rachel Cavill \\ Dr. Jerry Spanakis \\ Dr. Amrapali Zaveri (IDS) \\ Dr. Michel Dumontier (IDS)}
\date{March 1, 2019}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}

% Our images were all over the place, commenting this out seems to fix this mostly
%\usepackage{nonfloat}
% \usepackage{float}
\usepackage[export]{adjustbox}
\usepackage{spverbatim}
\usepackage{caption}
\usepackage{natbib}
\usepackage{scrextend}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{color}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage{wasysym}
\usepackage{changepage}
\usepackage{pgfgantt}
\usepackage{colortbl}
\usepackage{todonotes}
\usepackage[normalem]{ulem}
\useunder{\uline}{\ul}{}
%\usepackage[margin=1in]{geometry}
%\usepackage{multicol}
\usepackage{listings}
\lstset{numbers=left,numberstyle=\tiny,basicstyle=\ttfamily\scriptsize, tabsize=1, stringstyle=\small, framexleftmargin=1pt, captionpos=b, frame=single, breaklines=true}
\lstdefinestyle{rdf}{numberblanklines=true, morekeywords={}}
\lstdefinestyle{sparql}{numberblanklines=true, morekeywords={SELECT, WHERE, FILTER, GROUP, BY, IN, AS, SERVICE, GRAPH, CONSTRUCT, WITH}}
\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{5}


\begin{document}

\maketitle
%\addcontentsline{toc}{chapter}{\numberline{Abstract}}
\begin{abstract}
While there exists an abundance of open biomedical data, the lack of high quality metadata makes it challenging for others to find relevant datasets and to reuse them for another purpose. In particular, metadata are useful to understand the nature and provenance of the data. A common approach to improving the quality of metadata relies on expensive human curation, which itself is time consuming and also prone to error. Towards improving the quality of metadata we use scientific publications to automatically predict metadata key: value. For prediction, we use a Convolutional Neural Network (CNN) and a Bidirectional Long-short term memory network (BiLSTM). Additionally, we also perform a comparison with multi-label classification methods. The best results we get are from CNN. We focus our attention on the NCBI Disease Corpus. 
\end{abstract}

\renewcommand{\abstractname}{Acknowledgements}
%\addcontentsline{toc}{chapter}{\numberline{Acknowledgements}}
\begin{abstract}
First of all, I would like to thank my supervisor Dr. Amrapali Zaveri for introducing me to the problem of metadata quality, her guidance, constant encouragement and support, and invaluable help. Additionally, I would also like to thank Dr. Michel Dumontier for giving me the opportunity to work at IDS and his guidance. I would also like to thank Pedro Hernandez Serrano for helping me out with my doubts throughout the project. I would like to thank Dr. Rachel Cavill and Dr. Jerry Spanakis my supervisors at DKE for supervising me and providing valuable insights. I would also like to thank my friends Gaurav Maheshwari, Priyansh Trivedi and Yash Patel for their helpful comments. Last but not the least, I would like to thank my family who always provided me with perpetual love and support. 
 
 
\end{abstract}

\tableofcontents

\listoffigures
 
\listoftables




\input{chapters/introduction.tex}
\input{chapters/relatedwork.tex}
\input{chapters/dataset.tex}
\input{chapters/nn-method.tex}
%\input{chapters/data.tex}
\input{chapters/results.tex}
%\input{chapters/evaluation.tex}
\input{chapters/conclusion.tex}

\bibliographystyle{plain}
\bibliography{references}

\end{document}